Performance of two different ontological and data access methods

### 2014-07-03 cold2014/magnus.tex
cleaning up references
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@726 43ef1190-90fd-4dd7-a769-d0121f02b80e

The application present users with a web interface where they can
input values for any specification they might want to search for.  In
Figure~\ref{initialView} we can see how the application initially
looks for a user. Here the user can input values for the different
specifications, for instance fuel type or CO$_2$ emission. After a
user is done choosing specifications he can execute the actual car
model search. The application extracts the data from the form and
comprises it into another format that can easily be used later on. The
next step begins in the search module. Here it starts a thread for
each possible car model to execute each individual search. In the end
the user is presented with a set of valid models to choose from, which
contain the chosen specifications, e.g. the user have searched for a
model with diesel fuel, automatic transmission, four or more gears and
a maximum of six seats.

There are done some pre-computations to make the application a little
more effective. This includes extracting all the model names from both
ontologies, initiate all possible partially defined products as a
PartialCar object and creating some internal structure. These
pre-computations were done so that it would be possible to present the
user with the name of each car model, not just the URI. We will also
describe more in depth how the car model search is done later on.

@@

\subsection{Full factorial experiment}
A full factorial experiment is when one got $k$ factors and $n$ amount
of levels. This means that there is $n^k$ \emph{factorial designs}
which result in $n^k$ designs to execute.  In the thesis every
experiment will only have two levels, but the factors may vary. This
means that we will use a $2^k$ full factorial design for every
experiment. The experiments consist of $2^k$ combinations with $k$
factors. For instance if we have three factors where every factor got
two levels we get $2^3$ designs, also called \emph{runs}. This results
in a \emph{planning matrix} that shows the actual runs in the
experiment as tabulated in Table~\ref{designsspec}. This contains the
chosen factors and their levels.

%\begin{wraptable}
%\begin{center}
%    \begin{tabular}{ | l | l  l  l |}
%    \hline
%    {\bf Run} &  {\bf Fuel type} & {\bf Transmission} & {\bf Speed} \\ \hline
%	 1 & Diesel & Automatic & 100 \\ \hline
%	 2 & Petrol & Automatic & 100 \\ \hline
%	 3 & Diesel & Manual & 100 \\ \hline
%	 4 & Petrol & Manual & 100 \\ \hline
%	 5 & Diesel & Automatic & 200 \\ \hline
%	 6 & Petrol & Automatic & 200 \\ \hline
%	 7 & Diesel & Manual & 200 \\ \hline
%	 8 & Petrol & Manual &  200 \\ \hline
%    \end{tabular}
%\end{center}
%\caption{A filled planning matrix with 3 factors}\label{designsspec}
%\end{wraptable}

@@

\subsubsection{Experiment with ten factors}
The last experiment we did was a ten factor experiment. The goal here
was to see which factors would be deemed significant in a complete car
model search. With the results from this experiment we could discuss
several aspects of using the different ontologies.  Here we added 4
new factors in comparison to the six factor experiment. The newly
added factors were \emph{weight}, \emph{nr. of doors},
\emph{nr. of seats} and \emph{fuel consumption}. We used the same
approach here as in the previous experiments to calculate the average
value for the new factors. The average values can be seen in the
planning matrix in appendix~\ref{appB}. With 10 factors the amount of
runs were $2^{10}$, which made the planning matrix too big to show here.

@@

In the Figure~\ref{3factorGraph} we can see that only one significant
effects, the O factor. In the results data found in appendix
\ref{appD} we can see that the Configuration Ontology was almost two
seconds slower than the VSO/COO runs.  Three out of four CO runs kept
the response time around two seconds, but one run had to use
approximately 3.5 seconds which resulted in an average run time of
$\approx2.3$ seconds.  In a study done to research the amount of time
a user would be willing to wait, they found out that the tolerable
waiting time for any information retrieval was approximately 2
seconds.~\cite{waitTime} This means that with three factors we were
still within the time frame for a good user experience.

@@

Another factor that would influence the results is the amount of
values the search has to iterate over.  In Table~\ref{averages}, we
presented all the average values and the average number of values
present in the Daimler and Renault data. We can see that CO have
$0.54$ more fuel types on average to iterate over and $0.3$ more
transmissions. This is not a lot, but can account for some delay in
the response time.

>>>>

The application present users with a web interface where they can
input values for any specification they might want to search
for. There the user can input values for the different specifications,
for instance fuel type or CO$_2$ emission. After a user is done
choosing specifications he can execute the actual car model
search. The application extracts the data from the form and comprises
it into another format that can easily be used later on. The next step
begins in the search module. Here it starts a thread for each possible
car model to execute each individual search. In the end the user is
presented with a set of valid models to choose from, which contain the
chosen specifications, e.g. the user have searched for a model with
diesel fuel, automatic transmission, four or more gears and a maximum
of six seats.

Some pre-computations are done to make the application a little
more effective. This includes extracting all the model names from both
ontologies, initiate all possible partially defined products as a
PartialCar object and creating some internal structure. These
pre-computations were done so that it would be possible to present the
user with the name of each car model, not just the URI. We will also
describe more in depth how the car model search is done later on.

@@

\subsection{Full factorial experiment}
A full factorial experiment is when one got $k$ factors and $n$ amount
of levels. This means that there is $n^k$ \emph{factorial designs}
which result in $n^k$ designs to execute.  We have chosen only
two-level experiments for this research, but the factors may vary. This
means that we will use a $2^k$ full factorial design for every
experiment. The experiments consist of $2^k$ combinations with $k$
factors. For instance if we have three factors where every factor got
two levels we get $2^3$ designs, also called \emph{runs}. This results
in a \emph{planning matrix} that shows the actual runs in the
experiment.

@@

\subsubsection{Experiment with ten factors}
The last experiment we did was a ten factor experiment. The goal here
was to see which factors would be deemed significant in a complete car
model search. With the results from this experiment we could discuss
several aspects of using the different ontologies.  Here we added 4
new factors in comparison to the six factor experiment. The newly
added factors were \emph{weight}, \emph{nr. of doors},
\emph{nr. of seats} and \emph{fuel consumption}. We used the same
approach here as in the previous experiments to calculate the average
value for the new factors. % TODO: Explain more how the levels were set?
With 10 factors the amount of
runs were $2^{10}$, which made the planning matrix too big to show here.

@@

In the Figure~\ref{3factorGraph} we can see that only one significant
effects, the O factor. We saw in Table~\ref{3factorEffect} that the
Configuration Ontology was almost two seconds slower than the VSO/COO
runs.  Three out of four CO runs kept the response time around two
seconds, but one run had to use approximately 3.5 seconds which
resulted in an average run time of $\approx2.3$ seconds.  In a study
done to research the amount of time a user would be willing to wait,
they found out that the tolerable waiting time for any information
retrieval was approximately 2 seconds.~\cite{waitTime} This means that
with three factors we were still within the time frame for a good user
experience.

@@

Another factor that would influence the results is the amount of
values the search has to iterate over.  We have found that CO have
$0.54$ more fuel types on average to iterate over and $0.3$ more
transmissions. This is not a lot, but can account for some delay in
the response time.
